---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

devtools::load_all()
library(tmaptools)
library(tmap)
library(AOI)
library(dplyr)
library(DT)
library(sf)

bb = st_bbox(c(xmin =-119.94455, ymin = 34.40026, 
               xmax = -119.74297, ymax = 34.46455), 
        crs = 4269) %>% 
    st_as_sfc() %>% 
    as_Spatial()


bg = rosm::osm.raster(bb, type = 'osmgrayscale')

```

# NFHL: National Flood Hazard Layers

<!-- badges: start -->

[![Build Status](https://travis-ci.org/mikejohnson51/NFHL.svg?branch=master)](https://travis-ci.org/mikejohnson51/NFHL) [![experimental](http://badges.github.io/stability-badges/dist/experimental.svg)](http://github.com/badges/stability-badges)

<!-- badges: end -->

The goal of NFHL is to provide access to the [National Flood Hazards Layers](https://www.fema.gov/national-flood-hazard-layer-nfhl) using the [AOI](https://github.com/mikejohnson51/AOI) subsetting workflows. Some preliminary thoughts on makeing this data more 'serviceable' are also given.

## Installation

You can install the development version of NFHL from  [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
remotes::install_github("mikejohnson51/NFHL")

# load libraries

library(NFHL)
library(AOI)
library(dplyr)
```

## Basic Workflow

### Layer IDs
The NFHL offers data for `r nrow(nfhl_meta)` layers whose type and ID can be found with `nfhl_meta`:

```{r, eval = F}
nfhl_meta
```

```{r, echo = F}
DT::datatable(nfhl_meta, 
              options = list(paging = FALSE,
                             search = FALSE,
                               dom = 't'),
              rownames = F)
```

### LayerID descriptions

Some of the layer names are vaugue. The `describe_nfhl` function scrapes the metadata for a given layerID and provides the name, description and spatial extents. 

Lets learn more about layerID 14, 'Cross-Sections':

```{r}
nfhl_describe(14)
```

### Getting Spatial Data

The `nfhl_get` function extacts the spatial data from the NDHL layers for a given spatial extent. 

A general workflow using the AOI/NFHL packages is 

  1. defining an AOI extent  
  2. identifying a NFHL layerID  
  3. extracting the data  

```{r}
# Define an AOI around UCSB
AOI <- aoi_get(list("UCSB", 10, 10))

# Look at layerID 28 description
nfhl_describe(28)$Description

# Extract Flood Hazard Polygons and filter
floodhazard <- nfhl_get(AOI, 28) %>% 
  filter(SFHA_TF == "T")
```

```{r, echo = FALSE}
 tm_shape(bg) + tm_rgb() +
 tm_shape(floodhazard) +
  tm_fill(col = 'STUDY_TYP') +
  tm_borders() +
  tm_compass(type = "8star", position = c("right", "bottom")) +
  tm_layout(
          legend.outside = TRUE,
          legend.outside.position = 'bottom',
          legend.bg.alpha = 1)
```

## Examples

### NHD and Cross-sections

Before we learned that the NFHL offered cross-sectional information (ID: 14). Lets get this data for our UCSB AOI, and overlay it with hydrographic data from the National Hydrograhpy Datast (NHD) found with the [HydroData](https://github.com/mikejohnson51/HydroData) package.

```{r, message = F, warning = F}
nhd <- HydroData::findNHD(AOI)[[2]]
# Note that any sf object can be passed to nfhl_get
cs  <- nfhl_get(nhd, 14)
```

```{r echo = FALSE}
tm_shape(bg) + 
  tm_rgb() +
 tm_shape(cs) +
   tm_lines() +
  tm_shape(nhd) +
   tm_lines(col = "blue") +
  tm_compass(type = "8star", position = c("right", "bottom")) 
```

## "Linked" NHD and General Structures Data

Lets look at one final example where we extract the "General Structures" identified in the NFHL (ID: 24) within the UCSB AOI.

```{r}
nfhl_describe(24)$Description
stru <- nfhl_get(AOI, 24)
```


```{r echo = FALSE}
tm_shape(bg) + tm_rgb() +
  tm_shape(nhd) +
  tm_lines(col = "blue") +
 tm_shape(stru) +
  tm_lines(col = 'STRUCT_TYP', lwd = 5) +
  tm_compass(type = "8star", position = c("right", "bottom")) +
    tm_layout(
      legend.outside = TRUE,
      legend.outside.position = 'bottom',
          legend.height = .5,
          legend.position = c("left","bottom"),
          legend.bg.color = "white",
          legend.bg.alpha = 1)
```

For some, linking these general structures to the NHD might be useful. Here we'll find those that cross the NHD, and add the 'crosses' relationship.

```{r, warning=F,message=F}
xx = st_join(stru, 
        st_transform(nhd, st_crs(stru)),
        join = st_crosses,
        left = FALSE) %>% 
  st_drop_geometry() %>% 
  mutate(realtionship = "crosses") %>% 
  select(OBJECTID, realtionship, comid, STRUCT_TYP, LAYER)

dim(xx)
head(xx)
```

In total we find 43 structures that cross the NHD, realtion 1 in the above tibble tells us that bridge 2533589 crosses COMID 17595405.

## Generalize to URIs

Each of these features (Structure and COMID) can be resolved to a URI using the following patterns:

```{r}
fema_base = 'https://hazards.fema.gov/gis/nfhl/rest/services/public/NFHL/MapServer/'
usgs_base = 'https://labs.waterdata.usgs.gov/api/nldi/linked-data/comid/'

linked = xx %>% 
  mutate(str_uri = sprintf("%s%s/query?&objectIds=%s&outFields=*&f=geoJSON", fema_base, LAYER,OBJECTID),
         realtionship,
         comid_uri = paste0(usgs_base, comid)) %>% 
  select(str_uri, realtionship, comid_uri)

head(linked)
```

## Working with URIs

The basic URI stings can then be directly called to operate on and use the represented features!

### Basic Mapping

```{r, echo = F}
bg2 = raster::crop(bg, read_sf(linked$comid_uri[1]))
```

```{r}
tm_shape(bg2) + tm_rgb() +
  tm_shape(read_sf(linked$str_uri[1])) +
  tm_lines(col = "red", lwd = 5) +
 tm_shape(read_sf(linked$comid_uri[1])) +
  tm_lines(col = 'blue', lwd = 2) +
  tm_layout(title = read_sf(linked$str_uri[1])$STRUCT_TYP)
```

### Elevation at Intersection

We can use our resource URIs to call the respective spatial feautures, intersect, them, extract the coordinates or intersection, and use those to query the USGS elevation API:

```{r, warning=F, message=F}
system.time({
  ## Coordinates of Intersection
  coords <- st_coordinates(st_intersection(read_sf(linked$comid_uri[1]), 
                           read_sf(linked$str_uri[1])))
  
  # Query USGS elevation API
  url <- sprintf('http://ned.usgs.gov/epqs/pqs.php?x=%s&y=%s&units=Meters&output=json', coords[1], coords[2]) 
          
  elev = read_json(url) 
})
```

```{r}
(elev)
```

## Acknowdegements

[Mike Johnson](http://mikejohnson51.github.io) is a graducate student in geography working with [Keith Clarke](http://www.geog.ucsb.edu/~kclarke/). This work contributes to the NSF funded [Convergence Accelorator Program](https://nsf.gov/awardsearch/showAward?AWD_ID=1937099&HistoricalAwards=false) on [Urban Flooding](https://ufokn.github.io/UFOKN/).

This package is experimental and comes with no garuntee `r emo::ji("smile")`. Pull requests are welcome!!
