---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  tidy.opts=list(width.cutoff=60),
  tidy=TRUE
)

library(NFHL)
library(tmap)
library(AOI)
library(dplyr)
library(sf)
```

# NFHL: National Flood Hazard Layers

<!-- badges: start -->
[![experimental](http://badges.github.io/stability-badges/dist/experimental.svg)](http://github.com/badges/stability-badges)
[![Dependencies](https://img.shields.io/badge/dependencies-4/19-green?style=flat)](#)
[![R CMD Check](https://github.com/mikejohnson51/NFHL/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/mikejohnson51/NFHL/actions/workflows/R-CMD-check.yaml)
[![Website](https://github.com/mikejohnson51/NFHL/actions/workflows/pkgdown.yaml/badge.svg)](https://github.com/mikejohnson51/NFHL/actions/workflows/pkgdown.yaml)
<!-- badges: end -->

The `NFHL` package provides access data from FEMA's [National Flood Hazards Layers](https://www.fema.gov/national-flood-hazard-layer-nfhl). It integates with the [AOI](https://github.com/mikejohnson51/AOI) package for spatial subsetting  workflows and offers functionality to extract and query NFHL layers. 

## Installation

You can install the development version of `NFHL` from  [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
remotes::install_github("mikejohnson51/NFHL")

# load libraries

library(NFHL)
library(AOI)
library(dplyr)
```

## Basic Workflow

### Layer IDs
The NFHL contains data for `r nrow(nfhl_meta)` layers. Use `nfhl_meta` to view the list of available layer IDs and names.

```{r, eval = F}
nfhl_meta
```

```{r, echo = F}
print(nfhl_meta)
```

### Layer descriptions

To get detailed information, such as layer name, description, and bounding box, use the `nfhl_describe()` function. For example, let's explore Layer ID 14 (Cross-Sections):

```{r}
nfhl_describe(14)
```

### Extracting Spatial Data

The `nfhl_get()` function retrieves spatial data for a specific layer and area of interest (AOI). The general workflow involves:

1. Defining an AOI.
2. Selecting a layer ID.
3. Extracting spatial data.

```{r}
# Define an AOI around UCSB
AOI <- AOI::aoi_ext("UCSB", wh = 10, units = "km", bbox = TRUE) |> 
  st_as_sf()

# View at Layer ID 28 description
nfhl_describe(28)$Description

# Extract Flood Hazard Polygons and filter by Special Flood Hazard Areas (SFHA)
floodhazard <- nfhl_get(AOI, 28) %>% 
  filter(SFHA_TF == "T")
```

```{r, echo = FALSE}
 tm_shape(floodhazard) +
  tm_fill(col = 'STUDY_TYP') +
  tm_borders() +
  tm_compass(type = "8star", position = c("right", "bottom")) +
  tm_layout(
          legend.outside = TRUE,
          legend.outside.position = 'bottom',
          legend.bg.alpha = 1)

```

## Examples

### Overlaying NHD and NFHL Cross-Sections

Layer ID 14 provides cross-sectional data. Let’s overlay it with hydrographic data from the National Hydrography Dataset (NHD) using the `nhdplusTools` package:

Before we learned that the NFHL offered cross-sectional information (ID: 14). Lets get this data for our UCSB AOI, and overlay it with hydrographic data from NHD.

```{r, message = F, warning = F}
# Get NHD data for the AOI
nhd <- nhdplusTools::get_nhdplus(AOI)

# Extract NFHL Cross-Section data (ID: 14)
cs  <- nfhl_get(nhd, 14)
```

```{r echo = FALSE}
 tm_shape(cs) +
   tm_lines() +
   tm_shape(nhd) +
   tm_lines(col = "blue") +
   tm_compass(type = "8star", position = c("right", "bottom")) 
```

## Linking General Structures with NHD

Layer ID 24 provides data on "General Structures." Here, we extract this layer, overlay it with NHD data, and identify structures that cross NHD features.

```{r}
# Describe Layer ID 24
nfhl_describe(24)$Description

# Extract General Structures
stru <- nfhl_get(AOI, 24)

# Identify crossings between structures and NHD data
crossings = st_join(stru, 
             st_transform(nhd, st_crs(stru)),
             join = st_crosses,
             left = FALSE) %>% 
  st_drop_geometry() %>% 
  mutate(realtionship = "crosses") %>% 
  select(OBJECTID, realtionship, comid, STRUCT_TYP, LAYER)

# View results
head(crossings)
```

Here, we identified `r nrow(crossings)` structural crossing's in the AOI river networ. For example, bridge 932599 crosses COMID 17596113.

```{r echo = FALSE}
  tm_shape(nhd) +
  tm_lines(col = "blue") +
 tm_shape(stru) +
  tm_lines(col = 'STRUCT_TYP', lwd = 5) +
  tm_compass(type = "8star", position = c("right", "bottom")) +
    tm_layout(
      legend.outside = TRUE,
      legend.outside.position = 'bottom',
          legend.height = .5,
          legend.position = c("left","bottom"),
          legend.bg.color = "white",
          legend.bg.alpha = 1)
```

Here, we identified `r nrow(crossings)` structural crossing's in the AOI river networ. For example, bridge 932599 crosses COMID 17596113.

## Acknowledgements

[Mike Johnson](http://mikejohnson51.github.io) is the Geospatial Science and Technology Lead at NOAA's Office of Water Prediction (OWP). This work contributes to the NSFfunded [Convergence Accelorator Program](https://nsf.gov/awardsearch/showAward?AWD_ID=1937099&HistoricalAwards=false) on [Urban Flooding](https://ufokn.github.io/UFOKN/).

This package is experimental and comes with no guarantee. Pull requests are welcome!!
